---
title: "Reproducible Research Week 2 Project"
author: "Xin Ha"
date: "3/16/2020"
output: html_document
---

## Load libraries

```{r}
library(ggplot2)
```
## Load and process the data
#### 1. Unzip and load data

```{r}
unzip("repdata_data_activity.zip")
activity <- read.csv("activity.csv")
```
#### 2. Calculate total number of steps and rename columns

```{r}
StepsbyDay <- aggregate(activity$steps, by=list(activity$date), FUN=sum, na.rm=FALSE)
names(StepsbyDay)[1] <- "Date"
names(StepsbyDay)[2] <- "Steps"
```
#### 3. Disply Results

```{r}
head(StepsbyDay, 61)
```
#### 4. Plot historgram of total number of steps taken

```{r}
ggplot(StepsbyDay, aes(x = Steps)) +
  geom_histogram(fill = "Blue", binwidth = 1200) +
  labs(title = "Total Steps by Day", x = "Steps", y = "Frequency")
```

### 5. Calculate Mean and Median number of steps per day

```{r}
meansteps <- mean(StepsbyDay$Steps, na.rm=TRUE)
mediansteps <- median(StepsbyDay$Steps, na.rm=TRUE)
```
* Mean steps: `r meansteps`
* Median steps: `r mediansteps`

## Average daily activity pattern

```{r}
activityremoveNA <- subset(activity,activity$steps >= 0)
MeanStepsbyInterval <- aggregate(activityremoveNA$steps, by=list(activityremoveNA$interval), FUN=mean)
names(MeanStepsbyInterval)[1] ="Interval"
names(MeanStepsbyInterval)[2] ="Steps"
```

### 1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
ggplot(MeanStepsbyInterval, aes(x=Interval, y=Steps)) +
  geom_line(color="blue") +
  labs(title = "Mean of Steps by Interval", x = "Interval", y = "Steps")
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
maxInterval <- MeanStepsbyInterval[which.max(MeanStepsbyInterval$Steps),]
```
* Maximum mean interval (Interval, Steps): `r maxInterval`

## Missing Values

### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
numberofNA <- subset(activity,is.na(activity$steps))  
missingvalues <- nrow(numberofNA)
```
* Number of missing values: `r missingvalues`

### 2. Devise a strategy to fill in Missing values.
#### Use the mean of interval to impute missing values.

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
numberofNA <- numberofNA[,c('date','interval')]
numberofNA <- merge(numberofNA,MeanStepsbyInterval,by.x='interval',by.y='Interval')
names(numberofNA)[3] <- 'steps'
activity_imputeNA <- rbind(activityremoveNA,numberofNA)
activity_imputeNA <- activity_imputeNA[order(activity_imputeNA$date,activity_imputeNA$interval),]
```

### 4. Make a histogram of the total number of steps taken each day 

```{r}
StepsByDay2 <- aggregate(activity_imputeNA$steps, by=list(activity_imputeNA$date), FUN=sum, na.rm=FALSE)
names(StepsByDay2)[1] ="Date"
names(StepsByDay2)[2] ="Steps"
head(StepsByDay2,61)
ggplot(StepsByDay2, aes(x = Steps)) +
  geom_histogram(fill = "Blue", binwidth = 1200) +
  labs(title = "Total Steps by Day", x = "Steps", y = "Frequency")
```

### 5. Calculate Mean and Median number of steps per day

```{r}
meansteps2 <- mean(StepsByDay2$Steps, na.rm=TRUE)
mediansteps2 <- median(StepsByDay2$Steps, na.rm=TRUE)
```
* Mean steps with imputation: `r meansteps2`
* Median steps with imputation: `r mediansteps2`

### 6. Do these values differ from the estimates from the first part of the assignment?

Yes, the mean is the same but the median has risen by 1.19 steps.  

### 7. What is the impact of imputing missing data on the estimates of the total daily number of steps?

The effect of using mean data per interval as a data imputation method for missing values pushes overall data towards the mean.

## Are there differences in activity patterns between weekdays and weekends?

### 1. Create new variable that gives day of week and whether it is on weekend or not

```{r}
activity_imputeNA$date <- as.Date(activity_imputeNA$date)
activity_imputeNA$DayofWeek <- weekdays(activity_imputeNA$date)
activity_imputeNA$IsWeekend <- ifelse(activity_imputeNA$DayofWeek == "Saturday" | 
                                        activity_imputeNA$DayofWeek == "Sunday", "Weekend", "Weekday")
```

#### Display results

```{r}
head(activity_imputeNA)
```
### 2. Time series panel plot

```{r}
MeanbyIsWeekend <- aggregate(activity_imputeNA$steps, 
                             by=list(activity_imputeNA$IsWeekend, activity_imputeNA$interval), mean)
names(MeanbyIsWeekend)[1] ="IsWeekend"
names(MeanbyIsWeekend)[2] ="Interval"
names(MeanbyIsWeekend)[3] ="Steps"
ggplot(MeanbyIsWeekend, aes(x=Interval, y=Steps, color=IsWeekend)) +
  geom_line() +
  facet_grid(IsWeekend ~ .) +
  labs(title="Mean Steps by Interval by type of Day", x = "Interval", y = "Steps")
```

Yes, there are differences between weekdays and weekend. On average, there is less number of steps on weekends. 


